The Semantics and Implementation of \CEU: a Synchronous Reactive Language based on Esterel

### 2015-09-19 paper.tex
"Stack-Based Internal Events" como primeira subsecao de 2.

In the sections that follow, we review the main differences between \CEU and 
Esterel~\cite{ceu.sensys13}:
shared-memory concurrency (Section~\ref{sec.ceu.det}),
safe abortion with finalization (Section~\ref{sec.ceu.abrt}),
first-class timers (Section~\ref{sec.ceu.timers}),
and stack-based internal events (Section~\ref{sec.ceu.ints}).

@@

@@

\subsection{First-class Timers}
\label{sec.ceu.timers}

@@

\subsection{Internal Events}
\label{sec.ceu.ints}

Esterel makes no semantic distinctions between internal and external signals, 
both having only the notion of either presence or absence during the entire 
reaction~\cite{esterel.preemption}.
%
In \CEU, however, internal events follow a stack-based execution policy, 
similar to subroutine calls in typical programming 
languages~\cite{ceu.sensys13}, contrasting with queue-based event handling 
adopted for external events.
%
Figure~\ref{lst.prints} illustrates the use of internal signals (events) in 
Esterel and \CEU.
%
In the version in Esterel, on the occurrence of \code{A}, \code{B} is emitted 
and they both become active, resulting in the invocation of \code{f()} and 
\code{g()} in no particular order.
%
In the version in \CEU, the occurrence of \code{A} makes the program behave as 
follows (with the stack contents in italics):
%
{\small
\begin{enumerate}
\setlength{\itemsep}{0pt}
\item 1st trail awakes (line 5), emits \code{b}, and pauses.\\
    \emph{stack: [1st]}
\item 2nd trail awakes (line 9), calls \code{\_f()}, and terminates.\\
    \emph{stack: [1st]}
\item 1st trail (on top of the stack) resumes, calls \code{\_g()}, and 
    terminates.\\
    \emph{stack: []}
\item Both trails have terminated, so the \code{par/and} rejoins, and the 
program also terminates;
\end{enumerate}
}

>>>>

In the sections that follow, we review the main differences between \CEU and 
Esterel~\cite{ceu.sensys13}:
stack-based internal events (Section~\ref{sec.ceu.ints}),
shared-memory concurrency (Section~\ref{sec.ceu.det}),
safe abortion with finalization (Section~\ref{sec.ceu.abrt}),
and first-class timers (Section~\ref{sec.ceu.timers}).

@@

\subsection{Stack-Based Internal Events}
\label{sec.ceu.ints}

Esterel makes no semantic distinctions between internal and external signals, 
both having only the notion of either presence or absence during the entire 
reaction~\cite{esterel.preemption}.
%
In \CEU, however, internal events follow a stack-based execution policy, 
similar to subroutine calls in typical programming 
languages~\cite{ceu.sensys13}, contrasting with queue-based event handling 
adopted for external events.
%
Figure~\ref{lst.prints} illustrates the use of internal signals (events) in 
Esterel and \CEU.
%
In the version in Esterel, on the occurrence of \code{A}, \code{B} is emitted 
and they both become active, resulting in the invocation of \code{f()} and 
\code{g()} in no particular order.
%
In the version in \CEU, the occurrence of \code{A} makes the program behave as 
follows (with the stack contents in italics):
%
{\small
\begin{enumerate}
\setlength{\itemsep}{0pt}
\item 1st trail awakes (line 5), emits \code{b}, and pauses.\\
    \emph{stack: [1st-trail]}
\item 2nd trail awakes (line 9), calls \code{\_f()}, and terminates.\\
    \emph{stack: [1st-trail]}
\item 1st trail (on top of the stack) resumes, calls \code{\_g()}, and 
    terminates.\\
    \emph{stack: []}
\item Both trails have terminated, so the \code{par/and} rejoins, and the 
program also terminates;
\end{enumerate}
}

@@

\subsection{First-Class Timers}
\label{sec.ceu.timers}

@@
